\input{setup}

\date{29.01.2020}

\begin{document}

\renewcommand{\enquote}[1]{\emph{``#1''}} % Cannot be done earlier

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
	\titlepage
	\doclicenseThis
\end{frame}

\section{Configuration File Formats}

\subsection{Definitions}

\begin{frame}
	\frametitle{Basic Definitions}
	The \intro[execution environment]{execution environment} is information outside the boundaries of each currently running process~\cite{corbato1971multics}.

	Controlling the execution environment is essential for configuration management~\cite{cons2002pan,huang2015confvalley}, testing~\cite{van2010automating,wang2009context}, and security~\cite{goldberg1996secure,schreuders2012towards,perkins2009automatically,liang2003isolated}.
\end{frame}

\begin{frame}
	\frametitle{Configuration Setting}
	\input{../shared/def/configurationsetting.tex}
\end{frame}


\begin{frame}[fragile]
	\frametitle{Synonyms for Configuration Settings}
	\ExecuteMetaData[../book/background.tex]{synonyms}
\end{frame}

\begin{frame}[fragile]
	\frametitle{Definition}
	A \intro{configuration file} is a file containing configuration settings.

	\pause
	A Web server configuration file:

	\begin{lstlisting}[gobble=4]
	port=80 ; comment
	address=127.0.0.1\end{lstlisting}

	\only<2-2>{
	\begin{task}
	What are keys? What are configuration values? What is metadata?
	\end{task}
	}
	\pause

	The configuration values are ^80^ and ^127.0.0.1^, respectively.
	Other information in the configuration file is metadata for the configuration settings (such as the comment).
\end{frame}

\subsection{Formats}

\begin{frame}
	\frametitle{Types of Formats}
	\begin{itemize}
	\item CSV (comma-separated values)
	\item semi-structured
	\item programming language
	\item document-oriented
	\item literate
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{CSV formats}
	\begin{itemize}
	\item passwd: \formatdate{3}{11}{1971}
	\item passwd and group use : as separator
	\item are difficult to extend (e.g., GECOS)
	\item today mostly used for legacy reasons
	\item are replaced one-by-one (e.g., inetd, crontab)
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Trends}
	\begin{itemize}
	\item away from CSV
	\item towards general-purpose serialization formats (INI, JSON)
	\item human-read/writable (YAML, HOCON, TOML)
	\item programming language as configuration file
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Programming Language}
	\begin{description}
	\item[$+$] very easy for developers (simply source the file)
	\item[$+$] above-overage quality of error message
	\item[$-$] makes automatic change of individual values harder
	\item[$-$] very hard to use for people who do not know the programming language
	\item[$-$] does not separate code and data
	\end{description}
\end{frame}

\begin{assignment}
	\frametitle{Introduce somebody}
	\begin{task}
	Talk with someone about your favourite configuration file format.
	\end{task}

	\begin{task}
	Did you implement a configuration file parser and/or invented a new configuration file format?
	\end{task}

	\begin{task}
	Explain to everyone about the other person and his/her favourite configuration file format.
	\end{task}
\end{assignment}

\begin{frame}
	\frametitle{Method}

	What do FLOSS developers say?

	\begin{description}
	\item[\methodQuestion{}] survey with 672 persons visiting, 162 persons completing the survey~\cite{raab2017challenges}
	\item[\methodSource{}] source code analysis of 16 applications, comprising 50 million lines of code~\cite{raab2017challenges}
	\end{description}
\end{frame}

\begin{frame}
	\frametitle{Why are so many formats present?}
	\methodQuestion{} \question{In which way have you used or contributed to the configuration system/library/API in your previously mentioned FLOSS project(s)?}~\cite{raab2017challenges}
	\begin{itemize}
	\item \p{19} persons ($n=251$) have introduced a configuration file format.
	\item \p{29} implemented a configuration file parser.
	\item \p{15} introduced a configuration system/library/API.
	\item \p{34} used external configuration access APIs.
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Multitude of Formats}
	\begin{itemize}
	\item on every system a multitude of (legacy) configuration file formats exist
	\item the number grows fast
	\item thus applications usually have to deal with some legacy formats
	\end{itemize}
	

	\begin{restatable}{requirement}{reqLegacy}
	A configuration library must be able to integrate (legacy) systems and must fully support (legacy) configuration files.%
	\label{req:legacy}
	\end{restatable}
\end{frame}

\appendix

\begin{frame}[allowframebreaks]
	\bibliographystyle{plainnat}
	\bibliography{../shared/elektra.bib}
\end{frame}

\end{document}

