%TODO: add cliffhanger with preview for next time

\input{setup}

\date{27.3.2019}

\begin{document}

\renewcommand{\enquote}[1]{\emph{``#1''}} % Cannot be done earlier

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
	\titlepage
	\doclicenseThis
\end{frame}

\begin{frame}
	Lecture is every week Wednesday 09:00 - 11:00.

	\begin{description}
		\item[06.03.2019:] {\color{gray}topic, teams}
		\item[13.03.2019:] {\color{gray}TISS registration, initial PR}
		\item[20.03.2019:] {\color{gray}other registrations, guest lecture}
		\item[27.03.2019:] {\color{red}PR for first issue done, second started, \\ HS: kleiner Schiffbau}
		\item[03.04.2019:] {\color{orange}first issue done, PR for second}
		\item[10.04.2019:] mid-term submission of exercises
		\item[08.05.2019:] (HS?)
		\item[15.05.2019:]
		\item[22.05.2019:]
		\item[29.05.2019:]
		\item[05.06.2019:] final submission of exercises
		\item[12.06.2019:]
		\item[19.06.2019:] last corrections of exercises
		\item[26.06.2019:] exam
	\end{description}
\end{frame}

\begin{frame}
	\frametitle{Popular Topics}
	\vspace{-0.55cm}
	\setlength{\columnsep}{-1.3cm}
	\raggedright
	\begin{multicols}{2}
	\begin{description}
	\item[14] tools
	\item[9] testability
	\item[9] code-generation
	\item[7] context-awareness
	\item[6] {\color{orange} specification}
	\item[6] misconfiguration
	\item[6] {\color{red} complexity reduction}
	\item[5] validation
	\item[5] points in time % (early detection)
	\item[5] error messages
	\item[5] auto-detection
	\item[4] user interface
	\item[4] introspection
	\item[4] design
	\item[4] cascading
	\item[4] architecture of access
	\item[3] {\color{red} configuration sources}
	\item[3] {\color{orange} config-less systems}
	\item[2] secure conf
	\item[2] {\color{orange} architectural decisions}
	\item[1] push vs.\ pull
	\item[1] infrastructure as code
	\item[1] full vs.\ partial
	\item[1] convention over conf %iguration
	\item[1] CI/CD
	\item[0] documentation
	\end{description}
	\end{multicols}
\end{frame}

\begin{assignment}
	\frametitle{Tasks for today}
	(until 27.03.2019 23:59)

	\begin{task}
	Description of homework as pull request in private repo.
	(Inside a folder for you, use GitHub name.)
	\end{task}

	\begin{task}
	Description of teamwork (which application, which CM tool) as pull request in private repo.
	(Inside a folder for your team.)
	\end{task}

	\begin{task}
	Fix at least one issue and write some text in at least one other issue.
	\end{task}
\end{assignment}

\begin{frame}
	\frametitle{Slides}
	\begin{itemize}
	\item slide numbering: 02a is after 02
	\item old slides in same repo: always check date
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Some misconfigurations}
	\begin{itemize}
	\item studycode is Studienkennzahl
	\item same name twice in TALKS.xml
	\item \dots
	\end{itemize}

	\begin{task}
	How did these misconfigurations happened?
	\end{task}
\end{frame}

\begin{assignment}
	\frametitle{Tasks for next week}
	(until 03.04.2019 23:59)

	\begin{task}
	Fix misconfigurations in private repo.
	\end{task}

	\begin{task}
	Fix feedback about homework/teamwork.
	Calculate complexity of your teamwork.
	\end{task}

	\begin{task}
	First issue done, PR for second issue and write some text in at least one other issue (if 5 issues are not yet assigned to you).
	\end{task}
\end{assignment}


\begin{frame}
	\frametitle{Recapitulation}

	\methodQuestion{} \question{Which configuration systems/libraries/APIs have you already used or would like to use in one of your FLOSS project(s)?}
	\pause
	\begin{itemize}
	\item command-line arguments (\p{92}, $n=222$)
	\item environment variables (\p{79}, $n=218$)
	\item configuration files (\p{74}, $n=218$))
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Talk}
	Miruna Orsa

	Using XML in the day to day work life
\end{frame}










\begin{frame}
	\frametitle{Decision Tree}
	\begin{itemize}
	\item configuration settings may depend on each other
	\item form a decision tree~\cite{reiser2009cvm,czarnecki2012cool}
	\item the decision tree is an instantiation of chosen configuration settings
	\item calculation only needs to consider instantiations which make a difference: \\
	essential configuration complexity~\cite{meinicke2016essential}
	\end{itemize}
\end{frame}

\subsection{Usage}

\begin{frame}[fragile]
	\frametitle{Harmful Defaults~\cite{xu2015hey}}
	\begin{itemize}
	\item Problem: Two major data losses on a dozen machines.
	\item Cause:
	Stayed with the default values of the data-path settings
	(e.g., ^dfs.name.dir^, ^dfs.data.dir^) which point to locations in ^/tmp^.
	Thus, after the machines reboot, data losses occur.
	``One of the common problems from users.'' (from Cloudera)
	\item up to \p{53} of misconfigurations is due to staying at defaults
	\item \p{17} to \p{48} of configuration issues are about difficulties in finding settings
	\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{Unnecessary Settings~\cite{xu2015hey}}
	\begin{itemize}
	\item Configuration Parameter: ^dfs.namenode.tolerate.heartbeat.multiplier^
	\item Developers' Discussion:
	Since we are not sure what is a good choice, how about making it
	configurable?
	We should add a configuration option for it. Even if it's unlikely to
	change, if someone does want to change it they'll thank us that they
	don't have to change the code/recompile to do so.
	\item Real-World Usage:
	\begin{itemize}
	\item No usage found by searching the entire mailing lists and Google.
	\item No usage reported in a survey of 15 Hadoop users in UCSD.
	\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Unnecessary Settings~\cite{xu2015hey}}
	\begin{itemize}
	\item \p{6} to \p{17} of settings set by majority
	\item up to \p{54} are seldom set
	\item up to \p{47} of numeric settings have no more than five distinct values
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Reduction}
	\methodQuestion{}
	\question{Why do you think configuration should be reduced?}
	\begin{itemize}
	\item to simplify code maintenance (\p{50}),
	\item to prevent errors and misconfiguration (\p{43}),
	\item to provide better user experience (\p{40}),
	\item \textbf{\question{I do not think it should be reduced} (\p{30})},
	\item because they prefer auto-detection (\p{29}) \\ (with a possibility to override configuration settings: \p{32}),
	\item \question{because use-cases which are rarely used should not be supported} (\p{13}),
	\item \question{never find time for this task} (\p{9}), and
	\item \question{because only standard use-cases should be supported} (\p{1})
	\end{itemize}
\end{frame}

\begin{frame}
	\begin{alertblock}{Question}
	How to specify reduction strategies of configuration settings?
	\end{alertblock}
	\pause
	\begin{exampleblock}{Answer}
	Configuration Specification
	\end{exampleblock}
\end{frame}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\nocite{raab2017introducing}

\appendix

\begin{frame}[allowframebreaks]
	\bibliographystyle{plainnat}
	\bibliography{../shared/elektra.bib}
\end{frame}

\end{document}


