\chapter{Acknowledgements}

\chapterprecishere{Physics isn't the most important thing. Love is.\par\raggedleft--- \textup{Richard Feynman}}

First of all thanks to all people involved in \elektra{}.
Many people contributed over the years.
Others helped by participating in user studies.
Special thanks to Kai-Uwe Behrmann and Harald Geyer, who are the pioneer users of \elektra{}.
We had many fruitful discussions about many topics concerning configuration.

Thanks to the reviewing committee for finding time.

Thanks to the staff of my institute who proofread many of my papers and helped in all aspects around research.
I want to give distinct gratitude to Franz Puntigam for guiding me through the work and for being highly responsive to all my concerns and questions.
I thank TU Wien for financial support.

Furthermore, thanks to the people for helping me with reviews and discussions.
In particular, reversely sorted by alphabet, I am grateful to
Uwe Zdun,
Tianyin Xu,
Stefan Winter,
Patrick Sabin,
Nedko Tantilov,
Milan Nosáľ,
Maike Löhndorf,
Katharina Spiel,
Jens Knoop,
Helmut Toplitzer,
Harald Geyer,
Gergö Barany,
Geraldine Fitzpatrick,
Franz Puntigam,
Erik Schnetter,
Elisabeth Raab-Steiner,
Christian Amsüss,
Avi Alkalay,
Aotani Tomoyuki,
Andreas P. Priesnitz,
Andreas Mohrhammer,
Andreas Krall,
Andreas Falkner,
and the anonymous reviewers.

I build upon the work of many people who contributed to free software, science, and society.
Without them I would not have been able to even start this work.
I do apologize to all the people I cannot mention here!

Last but not least, I am deeply grateful for the backing I got from my family and friends, especially Natalie Kukuczka and Elisabeth Raab.
