\documentclass{beamer}

\input{setup}

\date{9.3.2018}

\begin{document}

\renewcommand{\enquote}[1]{\emph{``#1''}} % Cannot be done earlier

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
	\titlepage
	\doclicenseThis
\end{frame}

\begin{assignment}
	\frametitle{Language of the Talk?}
	\begin{task}
	Hands up if you prefer German.
	\end{task}
	Unanimous preference of German required, otherwise English.
\end{assignment}

\section{Configuration File Formats}

\subsection{Definitions}

\begin{frame}
	\frametitle{Basic Definitions}
	The \intro[execution environment]{execution environment} is information outside the boundaries of each currently running process~\cite{corbato1971multics}.

	Controlling the execution environment is essential for configuration management~\cite{cons2002pan,huang2015confvalley}, testing~\cite{van2010automating,wang2009context}, and security~\cite{goldberg1996secure,schreuders2012towards,perkins2009automatically,liang2003isolated}.
\end{frame}

\begin{frame}
	\frametitle{Configuration Setting}
	\input{../shared/def/configurationsetting.tex}
\end{frame}


\begin{frame}[fragile]
	\frametitle{Synonyms}
\intro[user preferences|see{configuration setting}]{User preferences}~\cite{jin2014configurations} and \intro[customization|see{configuration setting}]{customization}~\cite{anderson2002researching} stress that users make the change although that might not always be the case.
\intro[variability point|see{configuration setting}]{Variability points}~\cite{gunther2012software,rhein2016variability,villela2014survey,van2001notion,nadi2014mining,mens2016taxonomy} aim at describing the capability of software to adapt its behavior.
\intro[derivation decision|see{configuration setting}]{Derivation decision}~\cite{software1993reuse,czarnecki2012cool} puts the decisions to make and not the result in focus.
\intro[configuration parameter|see{configuration setting}]{Configuration parameter}~\cite{yin2011empirical,anderson1994towards} is easily confused with other kinds of parameters.
\intro[configuration item|see{configuration setting}]{Configuration item}~\cite{anthony2009context} or \intro[configuration option|see{configuration setting}]{configuration option}~\cite{rabkin2011static,zhang2013automated,zhang2014configuration} are sometimes not applicable, for example, ``proxy option'', or ``language item''.
\intro[configuration data|see{configuration setting}]{Configuration data}~\cite{huang2015confvalley} is often used in the context of programmable gate arrays and has a different meaning in that domain.
\end{frame}

\begin{frame}[fragile]
	\frametitle{Definition}
	A \intro{configuration file} is a file containing configuration settings.

	\pause
	A Web server configuration file:

\begin{lstlisting}
port=80 ; comment
address=127.0.0.1\end{lstlisting}
	\pause

	Their configuration values are ^80^ and ^127.0.0.1^, respectively.
	Other information in the configuration file is metadata for the configuration settings (such the comment).
\end{frame}

\begin{frame}
	\frametitle{}
\end{frame}

\begin{assignment}
	\frametitle{Introduce somebody}
	\begin{task}
	Talk with someone else about your favourite configuration file format.
	\end{task}

	\begin{task}
	Explain to everyone about the other person and the favourite configuration file format.
	\end{task}
\end{assignment}

\section{Command-line Arguments}
\section{Environment Variables}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\nocite{raab2017introducing}

\appendix

\begin{frame}[allowframebreaks]
	\bibliographystyle{plainnat}
	\bibliography{../shared/elektra.bib}
\end{frame}

\end{document}


