\input{../book/setup}
\input{../shared/setup}
\input{../book/commandsetup}

\setcounter{chapter}{1} % chapters start with 0


\title{Configuration Management Languages}
\author{Markus Raab}


\begin{document}

\begin{titlingpage}
\maketitle
\doclicenseThis
\end{titlingpage}

\paragraph{Configuration management language}
is a relatively vaguely defined term---it is a language where some kind of configuration is specified.
In this reading text for the lecture Configuration Management, we will investigate different kinds of configuration management languages.

We investigated who already created configuration specification languages to improve configuration access for configuration management tools, i.e. configuration management language, answering the following research question:
\begin{restatable}{RQ}{rqBackgroundSpecificationLanguages}
 \label{rq:background-specification-languages}
 Which configuration specification languages are suitable to improve configuration access for configuration management tools?
\end{restatable}
%\rqBackgroundSpecificationLanguages*

\begin{hypothesis}[RQ~\ref{rq:background-specification-languages}]
We expect to find a large variety of configuration specification languages that already improve configuration access for configuration management tools.
\end{hypothesis}

\section{Method}

We did a survey of all configuration specification languages as revealed by Google Scholar with the search term:
\begin{verbatim}
language
"configuration specification" OR
"configuration description" OR
"configuration definition" OR
"configuration declaration"
\end{verbatim}

This search yielded several thousand articles.
We grouped them by dates because of download limits:
\begin{verbatim}
1950-1998     946 articles
1999-2004     919 articles
2005-2007     786 articles
2008-2010     872 articles
2011-2012     723 articles
2013-2016     810+ articles
\end{verbatim}

The ^+^ sign means that we subscribed to the search term to keep track of new incoming articles.
We scanned through the titles of all papers---or if this was not enough, we read the abstract---to filter off-topic papers.
In particular, we removed all articles that describe general purpose languages, behavioral descriptions, or that are domain-specific.
After this process, we grouped papers that described the same configuration specification language.
As result, we found 92 configuration specification languages.
Due to lack of time, we only further processed the ones that are at least remotely related to \elektra{Spec}, the specification language of \elektra{}, and are of interest for this reading text.
In this step, we excluded about \sfrac{3}{4} of the configuration specification languages.

In the rest of the section, we will describe four selected properties, i.\,e.\ expressiveness, reasoning, modularity, and reusability, for some configuration specification languages.
Others are mentioned in ``Others''.



\section{CFEngine}

CFEngine~\cite{burgess1995cfengine,pandey2012investigating} is a language-based system administration tool that pioneered idempotent behavior.
It uses declarative class-based decision structures.
\citet{burgess2003theory}~introduces theory behind it.

\paragraph*{Expressiveness:}
CFEngine allows us to declare dependences and facilitates some high-level configuration specification constructs.
In its initial variants it neither had validation specifications, cardinalities, nor higher-level relationships.
\paragraph*{Reasoning:}
\notsupported{}
\paragraph*{Modularity:}
\notsupported{}

\paragraph*{Reusability:}
Existing system administrator scripts can be profitably run from CFEngine.


\section{NIX}

The NIX language~\cite{dolstra2007purely} claims to be purely functional as a novel feature.
The main concept is the referential transparency both for the configuration specification language and for the system itself.
A large-scale deployment shows that the approach is feasible and practical.

Read \url{https://nixos.org/} for more information.

\paragraph*{Expressiveness:}
NIX expressions, for example functions, describe how to build software packages.
The unit of variability is a package.
Additionally, a hierarchical set of properties describes the configuration specification.
Otherwise, the expressiveness is low, NIX describes neither cardinalities nor relationships.

\paragraph*{Reasoning:}
Because of the referential transparency of the system itself, every solution derived from the NIX expressions should be valid, so no reasoning or conflict handling is necessary.
Some operations, however, might lead to a completely new system.

\paragraph*{Modularity:}
The NIX expressions are modular because they ensure absence of side effects and thus can be easily composed.

\paragraph*{Reusability:}
Derivations that describe atomic build actions are reused in other derivations.
Import and inherit features are used to create packages, improving reusability.




\section{ConfValley (CPL)}

\citet{huang2015confvalley} introduce systematic validation for cloud services.
ConfValley uses a unified configuration settings representation for tens of different configuration file formats.
Its configuration specification language, called CPL, does not aim to be a type-safe configuration specification language.
It enables, however, system administrators of cloud services to write declarative specifications of properties with correctness constraints.

\paragraph*{Expressiveness:}
CPL introduces many concepts and has non-trivial language features.
Its most expressive elements are first-order quantifiers.
CPL is not able to specify dynamic and complex requirements.

\paragraph*{Reasoning:}
Constraints can be inferred by running an inference engine on configuration settings that are considered good (black-box approach).
Within the validation engine, however, no constraint solver is available.

\paragraph*{Modularity:}
CPL aims at easy grouping of constraints.
Its extensibility has limitations, for example, adding language primitives need modifications in the compiler.
The authors claim, however, that these changes can be done in a straightforward way---at least for predicates.%
{\parfillskip=0pt plus .7\textwidth \emergencystretch=.5\textwidth \par}

\paragraph*{Reusability:}
Using transformations and compositions, predicates can be reused in different contexts.
Also with language constructs like ^let^, specifications can be reused.




\section{Quattor (Pan)}

\citet{cons2002pan} invented and used PAN for many machines within CERN.
Furthermore, the language is still used by Quattor.
The configuration database in Pan comprises high-level and low-level descriptions.
The low-level descriptions are in XML syntax.
Here we focus on the declarative, high-level description.

\paragraph*{Expressiveness:}
The Pan language allows users to specify data types, validation with code snippets and constraints.
It only supports lists but no configurable cardinality nor is-a/part-of relationships.
The compiler uses a 5 step process: compilation, execution, insertions-of-defaults, validation, and serialization.

\paragraph*{Reasoning:}
Pan focuses on validating configurations, it is not able to generate new configurations.
Pan provides type enforcement with embedded validation code.

\paragraph*{Modularity:}
The language has user-defined data types (called templates) but otherwise has only minimal support for modularity.
In particular, side effects and assignments hinder modularity of validation code.

\paragraph*{Reusability:}
Reusability and collaboration is only possible via simple include statements and a simple inheritance mechanism of templates.


\section{UML}

\citet{felfernig1999knowledge,felfernig2000uml,felfernig2002joint} describe an approach where the unified modeling language (UML) is used as notation to simplify the construction of a logic-based description.
The papers formally describe the semantics. Tools are available and experimental results show feasibility.

\paragraph*{Expressiveness:}
All UML features, including cardinality, domain-specific stereotypes and OCL-constraints are available.
The basic structure of the system is specified using classes, generalization and aggregation.
Resources impose additional constraints on the possible system structure.
Finally, the require-relation and incompatible-relation allow us to limit valid configurations.

\paragraph*{Reasoning:}
Customers provide additional input data and requirements for the actual variant of the product.
The logical sentences are range-restricted first-order-logic with a set extension and interpreted function symbols.
For decidability, the term-depth is limited to a fixed number.
It is possible to show that the configuration is consistent or that no solution exists.

\paragraph*{Modularity:}
Generalization is present without multiple inheritance with disjunctive semantics, i.\,e., only one of the given subtypes will be instantiated.

\paragraph*{Reusability:}
For shared aggregation additional ports are defined for a part.


\section{Others}

\textsc{Proteus}~\cite{tryggeseth1995modelling} shows the tight relation between software configuration management, like Git or Svn, and configuration specification languages.
\textsc{Proteus} combines both worlds in a powerful build system.
For example:

\begin{code}[basicstyle=\tiny,morekeywords={family,attributes,end,physical,default,classifications},gobble=4,language=]
family CalcProg
	attributes
		HOME : string default "/home/ask/proteus/test";
		workspace := HOME ++ "/calc/src/"; // string concatenation
		repository := "calc/";
		end
	physical
		main => "main.C";
		defs => "defs.h";
		exe => "calc.x" attributes workspace := HOME ++ "/calc/bin"; end
		classifications status := standard.derived; end;
	end
end
\end{code}

\citet{lock2005strider} invented Strider that supports modeling and analysis of complex systems.

ConfSolve~\cite{hewson2011modelling,hewson2012declarative} is a configuration specification language that is translated to a standard constraint programming language called MiniZinc.
Their focus is in finding configurations for machines and not to compute configuration settings.
ConfSolve generates Puppet code for deployment.

Many other configuration specification languages have been found during the survey~\cite{roll2003towards,pandey2012investigating,hill2011modeling,anderson2002lcfg,deliverable1996tina,lujak2015orcas,sommerville1992configuration,giese2012industrial,huang2007system%
,novak2005automatic%Puppet, Quattor,..
,gunther2012software,berger2013survey,magableh2010primitive,friedrich1999consistency}, but they do not provide configuration access specifications for FLOSS applications.

\section{Result}

The result of the survey was that we could not find a configuration specification language to be used as basis for \elektra{Spec}.
Instead all configuration specification languages we investigated had a different focus, which leads us to our answer of:
\rqBackgroundSpecificationLanguages*

\begin{finding}
We have to reject our hypothesis for \rqref{background-specification-languages}:
We did not find any configuration specification language that supports our goal of improving configuration access for configuration management tools.
Instead earlier work had at least one of the following two assumptions:
\begin{itemize}
\item Configuration files need to be generated instead of directly accessing configuration settings.
\item Applications need to be reimplemented using new development methods.
Architecture description languages, software product lines, and similar approaches have this assumption.
\end{itemize}
\end{finding}
\par
Both assumptions hinder progress in improvements of configuration access for configuration management tools.

\bibliographystyle{plainnat}

\begingroup
\sloppy
\makeatletter
\g@addto@macro{\UrlBreaks}{\UrlOrds}
\makeatother
\bibliography{../shared/elektra.bib}
\endgroup



\end{document}
